---
title       : Commute Analysis
subtitle    : Trying to minimizing time on the road
author      : Roberto Congiu rcongiu@yahoo.com
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---


```{r echo=FALSE}
# read data
data <- read.csv("data.csv", header=TRUE,sep="\t" )
data$dow <- weekdays(as.Date(data$dt))

# find dates bound for data
dtf <- levels(data$dt)
dtfl <- sort(dtf)

 isWeekend <- Vectorize(function(s) {
  return( s == "Saturday" || s == "Sunday")
})

plotBarPlot <- function( d, title) {
  boxplot(travelDurationTraffic/60  ~ hour, 
        xlab = "Hour in the day", ylab = "Travel time (minutes)",
        data = d,main= title)
}

plotWeek <- function(data, title) {
  dayFactor = factor ( data$dow, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
# abbrevviate day names for labels
abbDays <- lapply( levels(dayFactor), function(s) {substring(s,0,3)} )
  
boxplot( travelDurationTraffic/60 ~ dayFactor, data = data, xlabel="Day of Week",
         names = abbDays,
         ylab="Travel time (minutes)", main = title)
}



fromhome <- data[data$route =="home-to-work",]
fromwork <- data[data$route =="work-to-home",]

weekdayh <- subset(fromhome ,  !isWeekend(dow) )
weekendh <- subset(fromhome ,  isWeekend(dow) )

weekdayw <- subset(fromwork ,  !isWeekend(dow) )
weekendw <- subset(fromwork ,  isWeekend(dow) )

```

## Commute from Pasadena to Santa Monica

* about 25 miles
* goes through two of the most congested Los Angeles Freeways, the 110 and the 10
* takes 30 minutes without traffic
* can take up to 2 hours on the bad day/time
* significally different depending on the day of the week
* we want to analyze both going to work, and going back home, and find optimal working hours
* use Bing API to sample traffic estimates every 5 minutes, for 3 weeks.


--- .class #id 

## Traffic by day of week

* we can see there's lots of variance on weekdays

```{r echo=FALSE , fig.height=10, fig.width=12}
par(mfrow=c(2,2))
plotWeek(fromhome,"Pasadena -> Santa Monica")
plotWeek(fromwork,"Santa Monica -> Pasadena")
```

---

## Let's see how it changes on a weekday


```{r echo=FALSE , fig.height=10, fig.width=12}
par(mfrow=c(2,2))
plotBarPlot(weekdayh,"Pasadena to Santa Monica")
plotBarPlot(weekdayw,"Santa Monica to Pasadena")
```

---

## how about the weekend ?

```{r echo=FALSE , fig.height=10, fig.width=12}
par(mfrow=c(2,2))
plotBarPlot(weekendh,"Pasadena to Santa Monica")
plotBarPlot(weekendw,"Santa Monica to Pasadena")
```

What about dynamically estimating my commute based on 
the day of the week and the direction ?


---
## You can do that on my shiny app!

https://rcongiu.shinyapps.io/commute/




